---
title: "buildProbabilisticClassifier_comb"
author: "Lincoln Harris"
date: "4/16"
output: html_document

### Workflow for building a RF classifier that spits out probabilities, rather than 
###   discreet cell type assignments, from Linnearson hippo data
###
### Make sure mRF_func.R and externalFuncs.R are in the environment
###
### 4.13. What happens if i create separate Seurat objects, combine them, then train 
### classifier on this combined object? 
###
### THIS WORKS!!!
---

read in hippo loom file
```{r}
hippo_loom <- connect(filename = "l1_hippocampus.loom", mode = "r+")
amyg_loom <- connect(filename = "l1_amygdala.loom", mode = "r+")
```

coerce into dfs
```{r}
data_mat_h <- hippo_loom$matrix[,]
class(data_mat_h)
data_mat_h <- as.data.frame(data_mat_h)
class(data_mat_h)
dim(data_mat_h)

data_mat_a <- amyg_loom$matrix[,]
class(data_mat_a)
data_mat_a <- as.data.frame(data_mat_a)
class(data_mat_a)
dim(data_mat_a)
```

need to transpose bc genes and cells are inverted for some reason
```{r}
data_mat_h <- t(data_mat_h)
data_mat_a <- t(data_mat_a)
dim(data_mat_h)
dim(data_mat_a)
```

get vectors for gene and cell names
```{r}
# gives me gene names
hippo_genes <- hippo_loom$row.attrs$Gene[]
amyg_genes <- amyg_loom$row.attrs$Gene[]
length(hippo_genes)
length(amyg_genes)

# gives me cell names 
hippo_cells <- hippo_loom$col.attrs$CellID[]
amyg_cells <- amyg_loom$col.attrs$CellID[]
length(hippo_cells)
length(amyg_cells)

# setting row and col names for raw dfs
row.names(data_mat_h) <- hippo_genes
colnames(data_mat_h) <- hippo_cells

row.names(data_mat_a) <- amyg_genes
colnames(data_mat_a) <- amyg_cells
```

do we have dups in our cell names? 
```{r}
h_unique <- unique(hippo_cells)
a_unique <- unique(amyg_cells)

# also need a list of the duplicated elements
h_dups_index <- duplicated(hippo_cells)
a_dups_index <- duplicated(amyg_cells)

h_dups_cells <- hippo_cells[h_dups_index]
a_dups_cells <- amyg_cells[a_dups_index]

length(hippo_cells)
length(amyg_cells)

length(h_unique)
length(a_unique)

length(h_dups_cells)
length(a_dups_cells)
```

what about duplicate gene names? 
```{r}
length(unique(hippo_genes))
length(unique(amyg_genes))

h_dups_genes_index <- duplicated(hippo_genes)
a_dups_genes_index <- duplicated(amyg_genes)

h_dups_genes <- hippo_genes[h_dups_genes_index]
a_dups_genes <- amyg_genes[a_dups_genes_index]

length(h_dups_genes)
length(a_dups_genes)
```


Want to remove dups (cells AND genes) from raw data
```{r}
# get list of cells without ANY of the duplicated cells
hippo_cells_no_dups <- hippo_cells[!(hippo_cells %in% h_dups_cells)]
amyg_cells_no_dups <- amyg_cells[!(amyg_cells %in% a_dups_cells)]

# get list of genes without ANY of the duplicated ones
hippo_genes_no_dups <- hippo_genes[!(hippo_genes %in% h_dups_genes)]
amyg_genes_no_dups <- amyg_genes[!(amyg_genes %in% a_dups_genes)]

data_mat_h_no_dups <- data_mat_h[hippo_genes_no_dups, hippo_cells_no_dups]
data_mat_a_no_dups <- data_mat_a[amyg_genes_no_dups, amyg_cells_no_dups]

dim(data_mat_h_no_dups)
dim(data_mat_a_no_dups)
```

get cell_type from loom file metadata
```{r}
unique(hippo_loom$col.attrs$Class[])
h_cell_type <- hippo_loom$col.attrs$Class[]
h_subclass <- hippo_loom$col.attrs$Subclass[]

unique(amyg_loom$col.attrs$Class[])
a_cell_type <- amyg_loom$col.attrs$Class[]
a_subclass <- amyg_loom$col.attrs$Subclass[]

length(h_cell_type)
length(h_subclass)
length(a_cell_type)
length(a_subclass)
```

make my own metadata df
  these guys need dims of the initial (+ dups) dfs
```{r}
h_meta <- data.frame(matrix(nrow = 29519, ncol = 0))
h_meta$cell_names <- hippo_cells
h_meta$cell_type <- h_cell_type
h_meta$subclass <- h_subclass

a_meta <- data.frame(matrix(nrow = 25414, ncol = 0))
a_meta$cell_names <- amyg_cells
a_meta$cell_type <- a_cell_type
a_meta$subclass <- a_subclass
```

remove dups from metadata
```{r}
dim(h_meta)
dim(a_meta)

# need to do this instead of standard subsetting (ie. df2 <- df1[,cellsKeep])
#   bc we dont have rownames yet
h_meta_no_dups <- h_meta[!(h_meta$cell_names %in% h_dups_cells), ]
a_meta_no_dups <- a_meta[!(a_meta$cell_names %in% a_dups_cells), ]

dim(h_meta_no_dups)
dim(a_meta_no_dups)

#and now add the rownames
row.names(h_meta_no_dups) <- hippo_cells_no_dups
h_meta_no_dups$cell_names <- NULL

row.names(a_meta_no_dups) <- amyg_cells_no_dups
a_meta_no_dups$cell_names <- NULL
```

add metadata field so we can keep track of where all of the cells came from
```{r}
h_meta_no_dups$origin <- "hippo"
a_meta_no_dups$origin <- "amyg"
```

//////////////////////////////////////////////////////////////////////////
//////////////////////////////////////////////////////////////////////////
//////////////////////////////////////////////////////////////////////////

Create Seurat Objects
```{r}
hippo <- CreateSeuratObject(raw.data = data_mat_h_no_dups, meta.data = h_meta_no_dups)
amyg <- CreateSeuratObject(raw.data = data_mat_a_no_dups, meta.data = a_meta_no_dups)
```

MERGE SEURAT OBJECTS
  is this the right place to do the merge? 
```{r}
merged <- MergeSeurat(hippo, amyg, do.normalize = T, do.scale = F, do.center = F)
```

//////////////////////////////////////////////////////////////////////////
//////////////////////////////////////////////////////////////////////////
//////////////////////////////////////////////////////////////////////////

STANDARD SEURAT WORKFLOW

  not sure if i need to do this here? 
Seurat looks for UMI values in the gene expression matrix, for some reason. We need to tell it to look for n reads instead 
```{r}
colnames(merged@meta.data)[colnames(merged@meta.data) == 'nUMI'] <- 'nReads'
```

Find any ribosomal associated gene
```{r}
ribo.genes <- grep(pattern = "^Rp[sl][[:digit:]]", x = rownames(x = merged@data), value = TRUE)
```

Calculate percent ribo genes
```{r}
percent.ribo <- Matrix::colSums(merged@raw.data[ribo.genes,])/Matrix::colSums(merged@raw.data)
```

Add a percent ribo gene col to metadata (now a part of seurat obj)
```{r}
merged <- AddMetaData(object = merged, metadata = percent.ribo, col.name = "percent.ribo")
```

Sanity check: plot genes per cell vs reads per cell.
```{r}
GenePlot(object = merged, gene1 = "nReads", gene2 = "nGene", use.raw=T)
```

Filter cells by expression/gene count low/high thresholds
  what if we skip this step? 
```{r}
#merged <- FilterCells(object = merged, subset.names = c("nGene", "nReads"), low.thresholds = c(500, 50000), high.thresholds = c(25000, 2000000))
```

Normalize and scale
```{r}
# normalization has already been done
#merged <- NormalizeData(object = merged)
merged <- ScaleData(object = merged)
# predicting failure here
merged <- FindVariableGenes(object = merged, do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5)
merged <- RunPCA(object = merged, do.print = FALSE)
merged <- ProjectPCA(object = merged, do.print = FALSE)
```

running checks here
```{r}
PCElbowPlot(object = merged)
```

Find Clusters (Can play around with k.param. Default is 30. Decrease for smaller number of cells)
```{r}
merged <- FindClusters(object = merged, reduction.type = "pca", dims.use = 1:10, resolution = 0.5, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)
```

Running the actual tSNE here
```{r}
merged <- RunTSNE(object = merged, dims.use = 1:10, seed.use = 10)
```

…and visualize
```{r}
TSNEPlot(object = merged, do.label = T)
TSNEPlot(object = merged, do.label = T, group.by = "origin")
```

//////////////////////////////////////////////////////////////////////////
//////////////////////////////////////////////////////////////////////////
//////////////////////////////////////////////////////////////////////////

BUILD CLASSIFIER
  make sure external funcs are in the environment
```{r}
mergedClass_prob <- myRF_func(merged, training.classes = merged@meta.data$subclass, training.genes = row.names(merged@data))
```

//////////////////////////////////////////////////////////////////////////
//////////////////////////////////////////////////////////////////////////
//////////////////////////////////////////////////////////////////////////
